\documentclass{report}

\input{preamble}
\input{macros}
\input{letterfonts}



\title{\Huge{Math 2550}}
\author{\huge{PSET 2}}
\date{September 11 2025}

\begin{document}

\maketitle
\newpage% or \cleardoublepage
% \pdfbookmark[<level>]{<title>}{<dest>}
\pdfbookmark[section]{\contentsname}{toc}
\tableofcontents
\pagebreak

\section*{Ex 2.1}
\addcontentsline{toc}{section}{Ex 2.1}
\qs{}{
  Let $F$ be any ordered field (for example $F = \bbQ$ or $F = \R$).
  Suppose $A \subset F$ has a supremum. Suppose $\eps \in F$
  with $\eps > 0$. 
  Prove that there exists some $x \in A$ such that $x > (\sup A) - \eps$.
}

\begin{proofWithHibiscus}
  $A$ having a supremum means that there is some $y \in F$ that is a least upper bound for 
  $A$ such that for any upper bound $y' \in F$, $y \leq y'$. Yet for all $x \in A$, $x \leq y$. \\

  Let us suppose that the statement is false. That would mean that $\forall x \in A$, $x \leq (\sup A) - \epsilon$, where $\epsilon > 0$. \\

  Letting $\sup A = y$, this would mean that $y - \epsilon$ is an upper bound of $A$, since for all $x \in A$, $x \leq y - \epsilon$. But because $\epsilon > 0$, we have $y - \epsilon < y$. \\

  This contradicts the statement that $y$ is a least upper bound of $A$, such that for all other upper bounds $y'$, $y \leq y'$. So there must exist form $x \in A$ such that $s > (\sup A) - \epsilon$ 
\end{proofWithHibiscus}


\section*{Ex 2.2, Rudin 1.5}
\addcontentsline{toc}{section}{Ex 2.2, Rudin 1.5}

\qs{}{
  Let $A \subset \R$ be nonempty and bounded below. Define
  \begin{equation}
  -A = \{-x\ \mid x \in A\} \, .  
  \end{equation}
  Prove that $\inf A = - \sup(-A)$.
}

\begin{proofWithHibiscus}
  The infimum of $A$ is a number $y \in \R$ such that $y$ is a lower bound for $A$ and for any other lower bound $y'$ of $A$ we have $y' \leq y$. The supremum of a set $B$ is a number $z \in \R$ such that $z$ is an upper bound for $B$ and for any other upper bound $z'$ of $B$ we have $z \leq z'$.  \\

  $-A$ is defined by multiplying each element of $A$ by $-1$. Suppose $x, x' \in A$ with $x \leq x'$. Multiplying both sides by $-1$ yields $-x \geq -x'$. So the original ordering of A is reversed in -A. Smaller elements of $A$ correspond to larger elements of $-A$, and larger elements of $A$ correspond to smaller elements of $-A$. \\

  In particular, if $y$ is a lower bound of $A$, i.e.\ $y \leq x$ for all $x \in A$, then multiplying by $-1$ gives $-x \leq -y$ for all $x \in A$, so $-y$ is an upper bound of $-A$. If $u$ is an upper bound of $-A$, i.e.\ $-x \leq u$ for all $x \in A$, then $x \geq -u$ for all $x \in A$, so $-u$ is a lower bound of $A$. \\

  So let $y = \inf A$. Then for all $x \in A$ we have $y \leq x$, which is equivalent to $-x \leq -y$ for all $x \in A$. So $-y$ is an upper bound of $-A$. \\

  Now suppose $u$ is any other upper bound of $-A$. Then for all $x \in A$ we have $-x \leq u$, which means $x \geq -u$. So $-u$ is a lower bound of $A$, and since $y$ is the greatest lower bound of $A$ we have $-u \leq y$. Rearranging gives $u \geq -y$. Therefore $-y$ is the least upper bound of $-A$, so $\sup(-A) = -y$. \\
  \[
  \inf A = -\sup(-A).
  \]
\end{proofWithHibiscus}

  

\section*{Ex 2.3}
\addcontentsline{toc}{section}{Ex 2.3}

\qs{}{
  Suppose $A, B \subset \R$ are both nonempty and bounded above.
  Define the set $A+B$ by
  \begin{equation}
    A + B = \{a+b \mid a \in A, b \in B\} \, .
  \end{equation}
  Prove that $\sup(A + B) = \sup A + \sup B$.
}

\begin{proofWithHibiscus}
  The supremum of a set $S$ is a number $y \in \R$ such that $y$ is an upper bound for $S$ and for any other upper bound $y'$ of $S$ we have $y \leq y'$. \\

  Let $q = \sup A$ and $p = \sup B$. Then for all $a \in A$, $a \leq q$, and for all $b \in B$, $b \leq p$. Adding these gives $a+b \leq q+p$ for all $a \in A$, $b \in B$. So $q+p$ is an upper bound of $A+B$. \\

  WTS that $q + p$ is a least upper bound. Let $\epsilon > 0$.  
  By definition of the supremum, for every $\epsilon > 0$ there exists $a \in A$ with $q - \tfrac{\epsilon}{2} < a \leq q$.  
  Similarly, since $p = \sup B$, there exists $b \in B$ with $p - \tfrac{\epsilon}{2} < b \leq p$.  
  Then
  \[
    a+b > (q-\tfrac{\epsilon}{2}) + (p-\tfrac{\epsilon}{2}) = q+p - \epsilon.
  \]
  
  So for every $\epsilon > 0$, there is an element $a+b \in A+B$ greater than $q+p - \epsilon$. So $q+p$ is the least upper bound of $A+B$. \\

  Therefore
  \[
    \sup(A+B) = \sup A + \sup B.
  \]
\end{proofWithHibiscus}




\section*{Ex 2.4}
\addcontentsline{toc}{section}{Ex 2.4}

\qs{}{
   
  In each of the following, $S$ is an ordered set, and $A \subset S$.
  Answer the following in each case, and prove your answers: 
  \begin{itemize}
  \item Is $A$ bounded above? 
  \item Does $A$ have a maximum element, and if so, what is it? 
  \item Does $A$ have a supremum in $S$, and if so, what is it?
  \end{itemize}
  \begin{enumerate}
    \item $S = \Z$, $A = \{ 2, 3 \}$.
    \item $S = \Q$, $A = \{ -\frac{2n}{5} \mid n \in \bbN \}$.
    \item $S = \Q$, $A = \{ -\frac{1}{n} \mid n \in \bbN \}$.
    \item $S = \Q$, $A = \{ \frac{1}{n} \mid n \in \bbN \}$.
    \item $S = \Q$, $A = \{ x \in \Q \mid 0 < x \le 1 \}$.
    \item $S = \Q$, $A = \{ x \in \Q \mid 0 < x < 1 \}$.
    % \item $S = \Q$, $A = \{ x \in \Q \mid 0 < x^3 < 2 \}$.
  \end{enumerate}

}


\begin{proofWithHibiscus}
  \begin{enumerate}
    \item $S = \Z$, $A = \{2,3\}$. 
      \begin{enumerate}
        \item $A$ is bounded above since $3$ is an upper bound, and any $u \ge 3$ also works.  
        \item $A$ has a maximum. The largest element is $3$.  
        \item The supremum in $\Z$ is $3$, because the maximum is also the least upper bound.  
      \end{enumerate}
  
    \item $S = \Q$, $A = \{-\tfrac{2n}{5} \mid n \in \N\} = \{-\tfrac25, -\tfrac45, -\tfrac65, \dots\}$.  
      \begin{enumerate}
        \item $A$ is bounded above. The element $-\tfrac25$ is an upper bound.  
        \item $A$ has a maximum. The largest element is $-\tfrac25$.  
        \item The supremum in $\Q$ is $-\tfrac25$, since the maximum is the least upper bound.  
      \end{enumerate}
  
    \item $S = \Q$, $A = \{-\tfrac{1}{n} \mid n \in \N\} = \{-1, -\tfrac12, -\tfrac13, \dots\}$.  
      \begin{enumerate}
        \item $A$ is bounded above, because every element is $\le 0$.  
        \item $A$ does not have a maximum. The elements increase toward $0$, but $0 \notin A$.  
        \item $A$ has a supremum, which is $0$. First, $0$ is an upper bound. If $y<0$, choose $n > \frac{-1}{y}$. Then $-\tfrac1n > y$, so $y$ is not an upper bound. So $0$ is the least upper bound.  
      \end{enumerate}
  
    \item $S = \Q$, $A = \{\tfrac{1}{n} \mid n \in \N\} = \{1, \tfrac12, \tfrac13, \dots\}$.  
      \begin{enumerate}
        \item $A$ is bounded above, because $1$ is an upper bound.  
        \item $A$ has a maximum. The largest element is $1$.  
        \item The supremum is $1$, since the maximum is always the supremum.  
      \end{enumerate}
  
    \item $S = \Q$, $A = \{x \in \Q \mid 0<x \le 1\}$.  
      \begin{enumerate}
        \item $A$ is bounded above, because $1$ is an upper bound.  
        \item $A$ has a maximum. The largest element is $1$.  
        \item The supremum in $\Q$ is $1$, since it is the maximum.  
      \end{enumerate}
  
    \item $S = \Q$, $A = \{x \in \Q \mid 0<x<1\}$.  
      \begin{enumerate}
        \item $A$ is bounded above, because $1$ is an upper bound.  
        \item $A$ does not have a maximum. For any $x<1$ in $A$, the rational $x'=\tfrac{x+1}{2}$ is still in $A$ and larger than $x$.  
        \item $A$ has a supremum, which is $1$. First, $1$ is an upper bound. If $y<1$, pick $n$ large and set $x=1-\tfrac1n$. Then $x \in A$ and $x>y$, so $y$ is not an upper bound. So $1$ is the least upper bound.  
      \end{enumerate}
  \end{enumerate}
  \end{proofWithHibiscus}
  

\section*{Ex 2.5}
\addcontentsline{toc}{section}{Ex 2.5}


\qs{}{
  Suppose $F$ is an ordered field and $x, y, z \in F$.
  \begin{enumerate}
      % \item Prove that $0<1$.
      \item Prove that if $x>0$, then $x^{-1}>0$.
      \item Prove that if $x>0$, then $y>z$ if and only if $xy>xz$.
      \item Recall the field $\mathbb{F}_3 =\{0,1,2\}$ which we discussed in class. Prove that there does not exist an order on $\mathbb{F}_3$ such that it is an ordered field.
  \end{enumerate}
}

\begin{proofWithHibiscus}
  1. Let $F$ be an ordered field and let $x>0$. Because $F$ is a field, there exists
  $x^{-1}\in F$ with $x\cdot x^{-1}=1$, and $x^{-1}\neq 0$.
  Assume for contradiction that $x^{-1}\le 0$. Since $x^{-1}\neq 0$, we have $x^{-1}<0$.
  In an ordered field, multiplying an inequality by a negative number reverses the
  inequality; hence from $x>0$ we obtain
  \[
  x\cdot x^{-1} \;<\; 0 \quad\Longrightarrow\quad 1<0.
  \]
  But in any ordered field every nonzero square is positive, so $1=1^2>0$, a contradiction.
  Therefore $x^{-1}>0$. \\

  2. Let $x>0$. First suppose for contradiction that $y<z$ but $xy > xz$.  
  Since $x^{-1}>0$ by part (1), multiplying both sides of $xy > xz$ by $x^{-1}$ preserves the inequality:
  \[
  xy > xz \;\;\Rightarrow\;\; x^{-1}(xy) > x^{-1}(xz) 
  \;\;\Rightarrow\;\; (x \cdot x^{-1})y > (x \cdot x^{-1})z
  \;\;\Rightarrow\;\; y > z,
  \]
  contradicting $y<z$.
  
  For the other direction, suppose for contradiction that $y>z$ but $xy < xz$. Multiplying both sides by $x^{-1}>0$ preserves the inequality:
  \[
  xy < xz \;\;\Rightarrow\;\; x^{-1}(xy) < x^{-1}(xz)
  \;\;\Rightarrow\;\; (x \cdot x^{-1})y < (x \cdot x^{-1})z
  \;\;\Rightarrow\;\; y < z,
  \]
  contradicting $y>z$.
  
  Hence $y>z \iff xy > xz$. \\

  3. Suppose for contradiction that there is an order on $\mathbb{F}_3=\{0,1,2\}$ 
  making it an ordered field. By definition, an ordered field is a field with an order 
  satisfying two axioms. First, for all $x,y,z \in F$ with $y<z$, we must have 
  $x+y<x+z$. Second, for all $x,y \in F$ with $x>0$ and $y>0$, we must have $xy>0$. \\
  
  
  Since $\mathbb{F}_3=\{0,1,2\}$ has three elements, any total order must be a linear arrangement of these three numbers. There are $3!=6$ possible such orderings.  We show that each leads to a contradiction.  \\
  
  Case 1: $0<1<2$.  
  From $1<2$, by axiom (1) with $x=1$, we get $2<0$ (since $2+1=0$ in $\mathbb{F}_3$), 
  contradicting $0<\cdots<2$. \\
  
  Case 2: $0<2<1$.  
  From $0<2$, by axiom (1) with $x=1$, we get $1<0$, contradicting $0<\cdots<1$.  \\
  
  Case 3: $1<0<2$.  
  From $1<0$, by axiom (1) with $x=1$, we get $2<1$, contradicting $1<\cdots<2$.  \\
  
  Case 4: $1<2<0$.  
  From $2<0$, by axiom (1) with $x=1$, we get $0<1$, contradicting $1<\cdots<0$.  \\
  
  Case 5: $2<0<1$.  
  From $0<1$, by axiom (1) with $x=1$, we get $1<2$, contradicting $2<\cdots<1$.  \\
  
  Case 6: $2<1<0$.  
  From $2<1$, by axiom (1) with $x=1$, we get $0<2$, contradicting $2<\cdots<0$.  \\
  
  In every possible ordering of $\F_{3}$ assuming such an order leads to a contradiction with
  the property that for all $x,y,z \in F$ with $y<z$, we must have $x+y<x+z$. Therefore no order exists on $\mathbb{F}_3$ that would make it an ordered field.

\end{proofWithHibiscus}


\section*{Ex 2.6}
\addcontentsline{toc}{section}{Ex 2.6}

\qs{}{
  This exercise is a reminder that the ``size'' of infinite sets can
  be a bit counterintuitive.

  \begin{enumerate}
    \item
  Suppose $a, b \in \R$, with $a < b$.
  Prove that there is a bijection from $(a,b)$ to $(0,1)$.

  \item Prove that there is a bijection from $(0,1)$ to $(1,\infty)$.

  \item Suppose $a,b \in \R$. 
  Prove that there is a bijection from $(a,\infty)$ to $(b,\infty)$.
  \end{enumerate}
}

\begin{proofWithHibiscus}
  1. Suppose $a,b \in \R$ with $a<b$. We want a bijection from $(a,b)$ to $(0,1)$. 
  I want to send $a$ to $0$ and $b$ to $1$, so 
  \[
    f(x) = \frac{x-a}{b-a}.
  \]
  Since $a<x<b$, we get $0<f(x)<1$, so $f$ maps into $(0,1)$. 
  If $f(x_1)=f(x_2)$ then $\frac{x_1-a}{b-a}=\frac{x_2-a}{b-a}$, which gives $x_1=x_2$, so $f$ is injective. 
  For surjectivity, if $y\in(0,1)$ then $x=a+y(b-a)$ lies in $(a,b)$ and satisfies $f(x)=y$. 
  So $f$ is a bijection. \\

  2. We want a bijection from $(0,1)$ to $(1,\infty)$. 
  Let 
  \[
    g(x) = \frac{1}{x}.
  \]
  If $0<x<1$ then $1<1/x<\infty$, so $g(x)\in(1,\infty)$. 
  If $g(x_1)=g(x_2)$ then $1/x_1=1/x_2$, so $x_1=x_2$, so this function is injective. 
  For surjectivity, given $y>1$ we take $x=1/y$, which lies in $(0,1)$ and satisfies $g(x)=y$. 
  So $g$ is a bijection. \\

  3. Suppose $a,b \in \R$. WTS there is a bijection from $(a,\infty)$ to $(b,\infty)$. 
  Both sets are unbounded intervals, and WTS if we can list out all the elements of one interval as a sequence, then we can relabel that list to get the other interval. 

  Consider $(a,\infty)$. For each natural number $n$ we can define
  \[
    x_n = a+n.
  \]
  This gives a sequence $(x_n)$ in $(a,\infty)$, and in fact every point of $(a,\infty)$ lies between some consecutive terms of this sequence. Similarly, we can build a sequence in $(b,\infty)$ by
  \[
    y_n = b+n.
  \]
  Again, every point of $(b,\infty)$ lies between some consecutive $y_n$’s. 

  The map $h:(a,\infty)\to(b,\infty)$ defined by
  \[
    h(x) = x+(b-a)
  \]
  lines these sequences up. We have $h(x_n)=y_n$ for every $n$. If $x>a$, then $h(x)>b$, so the map goes into $(b,\infty)$. If $h(x_1)=h(x_2)$, then $x_1+(b-a)=x_2+(b-a)$ so $x_1=x_2$, proving injectivity. And given any $y>b$, we can write $y=h(x)$ with $x=y-(b-a)>a$, so the map is also surjective. 

  Thus $h$ is a bijection from $(a,\infty)$ to $(b,\infty)$.
\end{proofWithHibiscus}



\end{document}