\documentclass{report}

\input{preamble}
\input{macros}
\input{letterfonts}



\title{\Huge{Math 2550}}
\author{\huge{PSET 1}}
\date{September 4 2025}

\begin{document}

\maketitle
\newpage% or \cleardoublepage
% \pdfbookmark[<level>]{<title>}{<dest>}
\pdfbookmark[section]{\contentsname}{toc}
\tableofcontents
\pagebreak

\section*{Ex 1.1; Rudin 1.3}
\addcontentsline{toc}{section}{Ex 1.1; Rudin 1.3}
\qs{}{
  Suppose $F$ is a field with $x,y,z \in F$. Prove carefully from the field axioms:
  \begin{enumerate}
    \item If $x \neq 0$ and $xy = xz$, then $y = z$.
    \item If $x \neq 0$ and $xy = x$, then $y = 1$.
    \item If $xy = 1$, then $x \neq 0$ and $y = x^{-1}$.
    \item If $x \neq 0$, then $(x^{-1})^{-1} = x$.
  \end{enumerate}
}

\begin{proofWithHibiscus}
  \begin{enumerate}
    \item For all $x \in F$ with $x \neq 0$, we have $x^{-1} \in F$ such that 
    $x \cdot x^{-1} = 1$. So this means that we can multiply $(xy)$ and $(yz)$ by the multiplicative inverse of $x$
    \[ xy = yz \;\Rightarrow\; x^{-1}(xy) = x^{-1}(yz) \]
    Also, fields are associative: $\forall x,y,z \in F,\; x(yz) = (xy)z$. So we can rewrite our expressions. And using the multiplicative inverse property simplify to 1. 
    \[ (x^{-1} \cdot x) y = (x^{-1} \cdot x) z 
    \Rightarrow\; 1 \cdot y = 1 \cdot z \]

    Since $\forall x \in F,\; 1 \cdot x = x$, it follows that $y = z$.

    \item For all $x \in F$ with $x \neq 0$, we have $x^{-1} \in F$ such that 
    $x \cdot x^{-1} = 1$. So this means that we can multiply $(xy)$ and $1$ by the multiplicative inverse of $x$
    \[xy = x \;\Rightarrow\; x^{-1}(xy) = x^{-1}(x) \] 
    Also, fields are associative: $\forall x,y,z \in F,\; x(yz) = (xy)z$. So we can rewrite our expressions. And using the multiplicative inverse property simplify to 1. 
    \[ (x^{-1} \cdot x) y = (x^{-1} \cdot x) \Rightarrow\; 1 \cdot y = 1\]

    Since $\forall x \in F,\; 1 \cdot x = x$, it follows that $y = 1$.

    \item For all $x \in F$ with $x \neq 0$, we have $x^{-1} \in F$ such that 
    $x \cdot x^{-1} = 1$. So this means that we can multiply $(xy)$ and $1$ by the multiplicative inverse of $x$
    \[
    xy = 1 \;\Rightarrow\; x^{-1}(xy) = x^{-1}(1)
    \]
    Also, fields are associative: $\forall x,y,z \in F,\; x(yz) = (xy)z$. So we can rewrite our expressions. And using the multiplicative inverse property simplify to 1.
    \[
    (x^{-1} \cdot x) y = (x^{-1} \cdot 1) \;\Rightarrow\; 1 \cdot y = x^{-1}
    \]

    \item Given the field axioms that for all $x \in F$ with $x \neq 0$, we have $x^{-1} \in F$ such that 
    $x \cdot x^{-1} = 1$.  we can take the multiplicative inverse of $\left( x^{-1}\right)^{-1}$, $x^{-1}$ and multiply both sides by it giving
    \[ \left(x^{-1}\right)^{-1} \left(x^{-1}\right) 
        = x\left(x^{-1}\right) \]
    \[ \Rightarrow 1 = x \cdot x^{-1} \]

    and given the deifntion for multiplicative inverses, since $x \cdot x^{-1} = 1$ then x is an inverse of $x^{-1}$ which 
    was the inverse of $\left(x^{-1}\right)^{-1}$ meaning $x = \left(x^{-1}\right)^{-1}$
  \end{enumerate}
\end{proofWithHibiscus}

\section*{Ex 1.2, Rudin 1.3}
\addcontentsline{toc}{section}{Ex 1.2, Rudin 1.3}

\qs{}{
  In this course we have not yet defined $\R$. For this exercise, all you need to know about
  $\R$ is that it is a field which contains the field $\Q$.

  Suppose $r \in \Q$, $r \neq 0$, and $x \in \R$, $x \notin \Q$.  Prove that $r+x \notin \Q$ 
  and $rx \notin \Q$.
}

\begin{proofWithHibiscus}
  We know that $\Q$ is a field and $\R$ is a field, for the sake of contradiction, assume that there exists some $k = r + x \in \Q$. $\Q$ is a field with an addition rule, so $\forall x,y \in \Q $ we have some $x + y \in \Q$ but we also have the rule that $\forall x \in Q $ we have $-x + x = 0$ so since $k = r + x $ and $k \in \Q$ we can add the negation of $r$, $-r$ to both sides giving us $-r + k = -r + r + x$ and since both $-r$ and $k$ are in $\Q$ there sum must also be in $\Q$ but due to the negation rule that $\forall x \in \Q \exists -x $ such that $ - x + x = 0$ we get $-r + k = x$ which implies that $x \in \Q$ which contradicts the statement that $x \notin \Q$.  \\

  for the second part $rx \notin \Q$, let us assume for the sake of contradiction that there exist some $n = r \cdot k \in Q$. Since $\Q$ is a field with a multiplicative rule, $\forall x,y \in \Q $ we have $xy \in \Q$ and $\Q$ also has the multiplicative inverse rule meaning $\forall x \in Q $ we have some $x^{-1}$ such that $x^{-1} \cdot x = 1$. so assuming $n \in Q$ and $n = rx$ we can multiply both sides by $r^{-1}$, since $n $ and $r^{-1}$ are both in meaning their product must also be. So $n \cdot r^{-1} = r^{-1} \cdot r \cdot x $ and using the multiplicative inverse rule $r \cdot r^{-1} = 1 $ meaning $r^{-1} \cdot n = x \cdot 1$ and in a field $\forall x \in Q$ we have $1 \cdot x = x$, so this means $n \cdot r^{-1} = x$, which implies that $x \in \Q$ which contradicts the statement that $x \notin \Q$.
\end{proofWithHibiscus}

\section*{Ex 1.3}
\addcontentsline{toc}{section}{Ex 1.3}

\qs{}{
  In this problem you may freely use without proof all the standard facts about 
  arithmetic of rational numbers and integers.

  \begin{enumerate}

  \item Prove that $\bbQ$ does not contain any $x$ with $x^2 = 3$. (For this part it may be useful to use the
  fundamental theorem of arithmetic, which you may use without proof.)

  \item Let $\bbQ(\sqrt{3})$ be the set of ordered pairs
  $(a,b)$ with $a, b \in \bbQ$.
  Informally, we think of each pair $(a,b)$
  as representing a number ``$a + b \sqrt{3}$.''
  We can equip $\bbQ(\sqrt{3})$ with addition and product laws as follows:
  \begin{gather}
    (a,b)  +  (a',b') = (a+a', b+b') \, , \\
    (a,b) \cdot (a',b') = (aa' + 3bb', ab' + ba') \, .
  \end{gather}
  (These rules are motivated by applying the usual rules of arithmetic to sums of the form ``$a + b \sqrt3$''.)

  Show that $\bbQ(\sqrt{3})$ can be made into a field, with these addition and product laws.
  (This means saying carefully what are the $0$ and $1$ elements, what is the negation law, 
  what is the inversion law, and 
  then proving that all the axioms of a field are satisfied.
  This will probably be the lengthiest part of your writeup.)

  \item In solving part (2), did you use the result of part (1)? If so, where?

  \item $\bbQ(\sqrt3)$ contains an element $3$, defined as $3 = 1 + 1 + 1$, where $1$ is the $1$ element
  you already defined in part (2). What is the element $3$? Prove that $\bbQ(\sqrt3)$ contains exactly two elements $x$ with $x^2 = 3$.

  \item We could similarly consider the set $\bbZ(\sqrt{3})$, defined as above except that
  now we require $a, b \in \Z$. Show that
  $\bbZ(\sqrt{3})$ cannot be made into a field with the addition and product laws above.
  \end{enumerate}
}



\begin{proofWithHibiscus}
  \begin{enumerate}
    \item For sake of contradiction, assume that $\sqrt{3} \in \Q$. Then we can express
    \[
    \sqrt{3} = \frac{p}{q}
    \]
    with integers $p,q$ such that they share no common factors other then 1. Squaring gives
    \[
    3 = \frac{p^{2}}{q^{2}} \quad\Rightarrow\quad 3 q^{2} = p^{2}.
    \]
    Hence $p^{2}$ is a multiple of $3$, so $p$ is a multiple of $3$ meaning that p can be expressed as multiple of 3, $p = 3k$. Substituting,
    \[
    3q^{2} = (3k)^2 \;\Rightarrow\; 3q^{2} = 9k^{2} \;\Rightarrow\; q^{2} = 3k^{2}.
    \]
    So $q$ is also a multiple of $3$, contradicting the statement that they have no common factors other then 1. Therefore $\sqrt{3}\notin \Q$.

    \item \textbf{Field structure of $\Q(\sqrt{3})$.}
    \begin{enumerate}
      \item  The zero element is $(0,0)$ since
      \[
      (a,b)+(0,0)=(a,b)\quad\text{for all }(a,b).
      \]
      \item The additive inverse of $(a,b)$ is $-(a,b)=(-a,-b)$, because based on our defintion for addition in $\Q(\sqrt{3})$
      \[
      (a,b)+(-a,-b) = (-a + a, -b + b) =(0,0).
      \]
  
      \item The unit element is $(1,0)$, since
      \[
      (1,0)\cdot(a,b)=(1\cdot a+3\cdot0\cdot b,\,1\cdot b+0\cdot a)=(a,b)
      \]
      and $(a,b)\cdot(1,0)=(a,b)$ for all $(a,b)$.
  
      \item multiplicative inverse for $(a,b)\neq(0,0)$, we multiply by $(a,-b)$ since we need to get $(1,0)$
      \[
      (a,b)\cdot(a,-b)=(a^2-3b^2,\,0).
      \]
      If $a^2-3b^2=0$, then $\frac{a}{b}=\sqrt{3}$, for $b\neq0$, which is impossible since in part 1 we showed $\sqrt{3}$ is not in the set $\Q$. If $b=0$, then $a=0$ would contradict $(a,b)\neq(0,0)$. So $a^2-3b^2\neq0$ for every nonzero $(a,b)$, and we can define
      \[
      (a,b)^{-1}=\left(\frac{a}{a^2-3b^2},\; \frac{-b}{a^2-3b^2}\right).
      \]
      Then
      \[
      (a,b)\cdot(a,b)^{-1}=(1,0),
      \]
    \end{enumerate}
    Fields axioms: 
    \begin{enumerate}
      \item For all $x,y \in \Q(\sqrt{3})$ we have $x + y$ = $y + x$ \\
        let $x=(a,b)$, $y=(c,d)$ with $a,b,c,d\in\Q$. Then
        \[
          x+y=(a+c,\,b+d)=(c+a,\,d+b)=y+x,
        \]
        using commutativity in $\Q$ for each coordinate.
      \item For all $x,y,z \in \Q(\sqrt{3})$ we have $(x+y) + z = x + (y + z)$ \\
        Let $x=(a,b)$, $y=(c,d)$, $z=(e,f)$. Then
        \[
          (x+y)+z=(a+c+e,\,b+d+f)=x+(y+z),
        \]
        by associativity in $\Q$ coordinatewise.
      \item For al $x \in \Q(\sqrt{3})$ we have $0 + x = x$ \\
        With $0=(0,0)$ and $x=(a,b)$,
        \[
          0+x=(0+a,\,0+b)=(a,b)=x.
        \]
      \item For all $x \in \Q(\sqrt{3})$ we have $ - x + x = 0$\\
        For $x=(a,b)$ we have $-x=(-a,-b)$, so
        \[
        (-x)+x=(-a+a,\,-b+b)=(0,0)=0.
        \]
      \item For all $x,y \in \Q(\sqrt{3})$ we have $xy = yx$ \\
        With $x=(a,b)$, $y=(c,d)$,
        \[
          xy=(ac+3bd,\,ad+bc)=(ca+3db,\,cb+da)=yx,
        \]
        using commutativity in $\Q$.
      \item For all $x,y,z\in \Q(\sqrt{3})$ we have $(xy)z = x(yz)$ \\
        Let $x=(a,b)$, $y=(c,d)$, $z=(e,f)$. 
        \[
          (xy)z=\big(ac+3bd,\,ad+bc\big)(e,f)
              =\big(ace+3bde+3adf+3bcf,\; acf+ade+bce+3bdf\big),
        \]
        and
        \[
          x(yz)=(a,b)\big(ce+3df,\,cf+de\big)
              =\big(ace+3adf+3bcf+3bde,\; acf+ade+bce+3bdf\big),
        \]
      \item For all $x \in \Q(\sqrt{3})$ we have 1 $1\cdot x = x$ \\
      With $1=(1,0)$ and $x=(a,b)$,
        \[
          1\cdot x=(1,0)(a,b)=(1\cdot a+3\cdot 0\cdot b,\,1\cdot b+0\cdot a)=(a,b)=x.
        \]
      \item For all $x \in \Q(\sqrt{3})$ with $x \neq 0$ we have $x \cdot x^{-1}$ = 1 \\
        For $x=(a,b)\neq(0,0)$, we use our defintion of the multiplicative inverse 
        \[
          x^{-1}=\left(\frac{a}{a^2-3b^2},\,\frac{-b}{a^2-3b^2}\right).
        \]
        Since $(a,b)(a,-b)=(a^2-3b^2,0)$ and part (1) implies $a^2-3b^2\neq 0$ for $x\neq 0$,
        \[
          x\cdot x^{-1}
          =(a,b)\left(\frac{a}{a^2-3b^2},\,\frac{-b}{a^2-3b^2}\right)
          =(1,0)=1.
        \]
      \item For all $x,y,z \in \Q(\sqrt{3})$ we have $x(yz) = xy + xz$ \\
        With $x=(a,b)$, $y=(c,d)$, $z=(e,f)$,
        \[
          x(y+z)=(a,b)(c+e,\,d+f)
                =\big(a(c+e)+3b(d+f),\, a(d+f)+b(c+e)\big)
        \]
        \[
                =(ac+3bd,\,ad+bc)+(ae+3bf,\,af+be)=xy+xz.
        \]
    \end{enumerate}
    \item I did use part (1). It came up when I was trying to write down the formula for the inverse.  
    To make sure the denominator $a^2 - 3b^2$ isn't zero, I argued that if it were, then $\frac{a}{b} = \sqrt{3}$ (as long as $b\neq 0$). But part (1) told me that can’t happen in $\Q$. And if $b=0$ we'd just be left with $a=0$, which makes the whole element $(a,b)$ the zero element. 

    \item The unit element, $1$ is $(1,0)$, so $3$ is $(3,0)$. Now suppose $(a,b)$ squares to $3$:  
    \[
      (a,b)^2 = (a^2+3b^2,\, 2ab) = (3,0).
    \]  
    We see that $2ab=0$. So either $a=0$ or $b=0$. If $a=0$, then the first coordinate is $3b^2 = 3$, so $b=\pm 1$. If $b=0$, then we'd need $a^2 = 3$, but that would force $a$ to be $\sqrt{3}$, which is not rational. So that case doesn't work.  So the only square roots of $3$ in this field are $(0,1)$ and $(0,-1)$.

    \item Let $u=(0,1)$. If $\Z(\sqrt{3})$ were a field, then $u$ would have an inverse $(c,d)$ with integers $c,d$. But  
    \[
      (0,1)\cdot(c,d) = (3d,\,c).
    \]  
    For this to be our unit element we need $3d=1$ and $c=0$. That’s impossible with $d\in\Z$. So $u$ has no inverse here. Since every nonzero element has to have an inverse in a field, $\Z(\sqrt{3})$ can't be a field.
  \end{enumerate}
  

\end{proofWithHibiscus}


\section*{Ex 1.4}
\addcontentsline{toc}{section}{Ex 1.4}
\qs{}{
   
  Let $A$ be a nonempty subset of an ordered set $S$. If $\alpha\in S$ and, for all $x\in A$, we have $\alpha\leq x$, then we call $\alpha$ a \emph{lower bound} for $A$. Similarly, if $\beta\in S$ and, for all $x \in A$, we have $x\leq \beta$, then we call $\beta$ an \emph{upper bound} for $A$.

  Suppose $\alpha$ is a lower bound of $A$ and
  $\beta$ is an upper bound of $A$. 
  \begin{enumerate}
    \item Prove that $\alpha \le \beta$.
    \item Note that in the statement of the problem we require $A$ to be nonempty. Would the statement $\alpha \le \beta$ still hold even
    if we allow $A$ to be empty? Explain why. In solving part (1), did you use the fact that $A$ is nonempty?
    If so, where did you use it?
  \end{enumerate}    

}


\begin{proofWithHibiscus}
  1. Let us suppose that $\alpha$ is a lower bound of $A$ and $\beta$ is an upper bound of $A$. $A$ is a subset of $S$ where $S$ is an ordered set, now let us take an arbitrary element $x_{1} \in A$ and then by definition $\alpha \leq x_{1}$  and since $\beta$ is an upper bound, then $x_{1} \leq \beta$ then since $A$ is a subset of an ordered set and the relation $\leq $ is transitive, then $\alpha \leq x_{1}$ and $x_{1} \leq \beta$ means $\alpha \leq \beta$  \\

  2. I did use the statement that $A$ was non-empty so that I could find an arbitrary $x$ to compare $\alpha$ and $\beta$ so that I could use the rule of transitivity to imply $\alpha \leq \beta$. If $A = \varnothing$, then every $\alpha \in S$ would count as a lower bound and every $\beta \in S$ would count as an upper bound, but they wouldn’t necessarily satisfy $\alpha \leq \beta$. For example, in $\mathbb{R}$ with $A = \varnothing$, both $\alpha = 1$ and $\beta = 0$ are bounds, but clearly $1 \nleq 0$. So the statement fails. 
\end{proofWithHibiscus}


\end{document}